Coupling PDEs on 3D-1D domains with Lagrange multipliers

### 2020-03-02 infsup_discrete_conformingLM.tex
Updates from Overleaf

\section{Finite element approximation}
In this section we consider the discretization of Problem 1 and 2 by means of the finite element method. We address two main challenges; first we aim to identify a suitable approximation space for the Lagrange multiplier and to analyze the stability of the discrete saddle point problem; second we aim to derive a stable discretization method that uses independent computational meshes for $\Omega$, $\Gamma$ and $\Lambda$. Let us introduce a shape-regular triangulation $\mathcal{T}^{\Omega}_h$ of $\Omega$ and an admissible partition $\mathcal{T}^{\Lambda}_{h}$ of $\Lambda$.
We analyze two different cases: the one in which the 3D mesh is conforming to the interface $\Gamma$, namely the set of the intersections of the 3D elements of $\mathcal{T}^{\Omega}_h$ with $\Gamma$ is constituted by facets of such elements, and the non conforming case, namely the interface $\Gamma$ cuts the mesh arbitrarily. The discrete equivalent of \eqref{eq:saddle-point} reads as finding $u_h\in X_h\subset X$, $\lambda_h\in Q_h\subset Q$ s.t.
\begin{eqnarray}\label{eq:saddle-point_discrete}
\begin{cases}
a(u_h,v_h)+b(v_h,\lambda_h)=c(v_h)\quad &\forall v_h\in X_h\\
b(u_h,\mu_h)=d(\mu_h) \quad &\forall \mu_h\in Q_h.
\end{cases}
\end{eqnarray}
Let $B_h: Q'_h \longrightarrow Q_h$ be the operator induced by $b$ such that $\langle B_h v_h,\mu _h\rangle_{Q'_h,Q_h} = b(v_h,\mu _h)$.
The well posedness of such problem is governed by the classical result reported below.

\begin{corollary}{\cite[Theorem 2.42]{MR2050138}}Let $a(\cdot, \cdot)$ and $b(\cdot, \cdot)$ be continuous bilinear forms. 
Problem \eqref{eq:saddle-point_discrete} is well-posed if and only if 
\begin{align}\label{BNB1_discrete}
& \exists \alpha_h >0 : && \inf_{u_h\in ker(B_h)}\sup_{v_h\in ker(B_h)} \frac{a(u_h,v_h)}{\|u_h\|_{X}\|v_h\|_{X}}\geq \alpha_h\,,
\\
\label{eq:infsup_discrete}
 &\exists \beta_h >0: &&\inf_{\mu_h\in Q_h}\sup_{v_h\in X_h} \frac{b(v_h,\mu_h)}{\|v_h\|_{X}\|\mu_h\|_{Q}}\geq \beta_h\, .
\end{align}
\end{corollary}
This corollary is the discrete counterpart of Theorem \ref{th:bnb} where at the discrete level condition \eqref{BNB1_discrete} implies both of \eqref{BNB1}. We remark that \eqref{eq:infsup_discrete} does not follow from the conformity of the finte element spaces and it must be analysed independetly of \eqref{eq:infsup}.  
Let us notice that for both problem 1 and problem 2 the bilinear form $a(\cdot, \cdot)$ is coercive as stated in Lemmas \eqref{lemma:prob1_coercivity} and \eqref{lemma:prob2_coercivity}. Consequently, \eqref{BNB1_discrete} is automatically satisfied, being $\alpha_h$ the coercivity constant.\\

\subsection{$\mathcal{T}^{\Omega}_h$ conforming to $\Gamma$}
 We first analyze the case in which the 3D mesh is conforming to the interface $\Gamma$. As conformity conditions between $\mathcal{T}^{\Omega}_h$ and $\mathcal{T}^{\Lambda}_h$ with $\Gamma$ , we require that the intersection of $\mathcal{T}^{\Omega}_h$ and $\Gamma$ is made of entire faces of elemets $K \in \mathcal{T}^{\Omega}_h$. Furthermore, we also set a restriction between $\mathcal{T}^{\Omega}_h$ and $\mathcal{T}^{\Lambda}_h$. We assume that $\Lambda$ is a piecewise linear manifold. We want that the intersection of $\Gamma$ with any orthogonal plane to $\Lambda$ that crosses $\Lambda$ at the internal nodes of $\mathcal{T}^{\Lambda}_h$, consists of entire edges of $\mathcal{T}^{\Omega}_h$.

\subsubsection{Problem 1}
We denote by $X_{h,0}^k(\Omega)\subset H^1_0(\Omega)$, with $k>0$, the conforming finite element space of continuous piecewise polynomials of degree $k$ defined on $\Omega$ satisfying homogeneous Dirichlet conditions on the boundary and by $X_{h,0}^k(\Lambda)\subset H^1_0(\Lambda)$ the space of continuous piecewise polynomials of degree $k$ defined on $\Lambda$, satisfying homogeneous Dirichlet conditions on $\Lambda \cap \partial \Omega$.

@@

\subsubsection{Problem 2}
We aim to find  $u_h \in X_{h,0}^k(\Omega) ,\ {\ud}_h \in X_{h,0}^k(\Lambda), \ {\ld}_h \in Q_h \subset H^{-\frac12}(\Lambda)$, such that
\begin{subequations}
\begin{align}\label{eq:prob2_discrete}
&(u_h,v_h)_{H^1(\Omega)} + ({\ud}_h,{\vd}_h)_{H^1(\Lambda), |{\cal D}|} 
+  \langle  \mtrace v_h -  {\vd}_h, {\ld}_h \rangle_{\Lambda, |\DD|} 
\\
\nonumber
&\qquad\qquad= (f,v_h)_{L^2(\Omega)} + (\avrd{g},{\vd}_h)_{L^2(\Lambda), |{\cal D}|}
\quad \forall v_h \in X_h(\Omega), \ {\vd}_h \in X_h(\Lambda)
\\
&  \langle \mtrace u_h - {\ud}_h, {\md}_h \rangle_{\Lambda,| \DD| } = 0
\quad \forall {\md}_h \in Q_h\,.
\end{align}
\end{subequations}

>>>>

\section{Finite element approximation}
In this section we consider the discretization of the Problems 3D-1D-2D and 3D-1D-1D by means of the finite element method. We address two main challenges; first we aim to identify a suitable approximation space for the Lagrange multiplier and to analyze the stability of the discrete saddle point problem; second we aim to derive a stable discretization method that uses independent computational meshes for $\Omega$, $\Gamma$ and $\Lambda$. Let us introduce a shape-regular triangulation $\mathcal{T}^{\Omega}_h$ of $\Omega$ and an admissible partition $\mathcal{T}^{\Lambda}_{h}$ of $\Lambda$.
We analyze two different cases: the one in which the 3D mesh is conforming to the interface $\Gamma$, namely the set of the intersections of the 3D elements of $\mathcal{T}^{\Omega}_h$ with $\Gamma$ is constituted by facets of such elements, and the non conforming case, namely the interface $\Gamma$ cuts the mesh arbitrarily. The discrete equivalent of \eqref{eq:saddle-point} reads as finding $u_h\in X_h\subset X$, $\lambda_h\in Q_h\subset Q$ s.t.
\begin{eqnarray}\label{eq:saddle-point_discrete}
\begin{cases}
a(u_h,v_h)+b(v_h,\lambda_h)=c(v_h)\quad &\forall v_h\in X_h\\
b(u_h,\mu_h)=d(\mu_h) \quad &\forall \mu_h\in Q_h.
\end{cases}
\end{eqnarray}
This discrete problem is well-posed if the \eqref{BNB1}-\eqref{BNB4} conditions applies to $X_h$ and $Q_h$. Since $X_h\subset X$ and $Q_h\subset Q$,  
\eqref{BNB1}-\eqref{BNB3} follows immediately and only the inf-sup condition needs consideration.  We summarize this in the Corollary below. 
\begin{corollary}{\cite[Theorem 2.42]{MR2050138}}\\
Let
$X_h \subset X$, $Q_h \subset Q$, $a(\cdot, \cdot)$ and $b(\cdot, \cdot)$ satisfy the conditions  \eqref{BNB1}-\eqref{BNB3} then  
the problem \eqref{eq:saddle-point_discrete} is well-posed if 
the discrete counterpart of \eqref{BNB4} is satisfied, ie., 
there exists a $\beta_h>0$ such that
\begin{align}\label{BNB1_discrete}
\label{eq:infsup_discrete}
 \sup_{v_h\in X_h} \frac{b(v_h,\mu_h)}{\|v_h\|_{X}}\geq \beta_h |\mu_h\|_{Q}\, .
\end{align}
\end{corollary}

\subsection{$\mathcal{T}^{\Omega}_h$ conforming to $\Gamma$}
 We first analyze the case in which the 3D mesh is conforming to the interface $\Gamma$. As conformity conditions between $\mathcal{T}^{\Omega}_h$ and $\mathcal{T}^{\Lambda}_h$ with $\Gamma$ , we require that the intersection of $\mathcal{T}^{\Omega}_h$ and $\Gamma$ is made of entire faces of elements $K \in \mathcal{T}^{\Omega}_h$. Furthermore, we also set a restriction between $\mathcal{T}^{\Omega}_h$ and $\mathcal{T}^{\Lambda}_h$. We assume that $\Lambda$ is a piecewise linear manifold. We want that the intersection of $\Gamma$ with any orthogonal plane to $\Lambda$ that crosses $\Lambda$ at the internal nodes of $\mathcal{T}^{\Lambda}_h$, consists of entire edges of $\mathcal{T}^{\Omega}_h$.

\subsubsection{Problem 3D-1D-2D}
We denote by $X_{h,0}^k(\Omega)\subset H^1_0(\Omega)$, with $k>0$, the conforming finite element space of continuous piecewise polynomials of degree $k$ defined on $\Omega$ satisfying homogeneous Dirichlet conditions on the boundary and by $X_{h,0}^k(\Lambda)\subset H^1_0(\Lambda)$ the space of continuous piecewise polynomials of degree $k$ defined on $\Lambda$, satisfying homogeneous Dirichlet conditions on $\Lambda \cap \partial \Omega$.

@@

\subsubsection{Problem 3D-1D-1D}
We aim to find  $u_h \in X_{h,0}^k(\Omega) ,\ {\ud}_h \in X_{h,0}^k(\Lambda), \ {\ld}_h \in Q_h \subset H^{-\frac12}(\Lambda)$, such that
\begin{subequations}
\begin{align}\label{eq:prob2_discrete}
&(u_h,v_h)_{H^1(\Omega)} + ({\ud}_h,{\vd}_h)_{H^1(\Lambda), |{\cal D}|} 
+  \langle  \mtrace v_h -  {\vd}_h, {\ld}_h \rangle_{\Lambda, |\DD|} 
\\
\nonumber
&\qquad\qquad= (f,v_h)_{L^2(\Omega)} + (\avrd{g},{\vd}_h)_{L^2(\Lambda), |{\cal D}|}
\quad \forall v_h \in X_h(\Omega), \ {\vd}_h \in X_h(\Lambda)
\\
&  \langle \mtrace u_h - {\ud}_h, {\md}_h \rangle_{\Lambda,| \DD| } = 0
\quad \forall {\md}_h \in Q_h\,.
\end{align}
\end{subequations}

### 2020-03-02 unfitted_prob2.tex
Updates from Overleaf

\def\patch{\omega _j}
\subsection{$\mathcal{T}^{\Omega}_h$ non conforming to $\Gamma$ and $\Lambda$} \label{sec:unfit2}
We analyze now the case in which the elements of the 3D mesh $\mathcal{T}^{\Omega}_h$ do not conform with the surface $\Gamma$ nor with $\Lambda$. In this case, the 3D-1D-1D
formulation, namely Problem 2, is more suitable. 
%{\color{red} Add more details, we can refer also to cutFEM (Burman, Massing etc.) explaining the limitation of that approach (network case for example).}\\
Therefore we focus on the analysis of this discretization approach for Problem 2 solely.

\subsubsection{Problem 2} We consider for the solutions $u_h$ and ${\ud}_h$ the spaces $X_h = X_{h,0}^1(\Omega) \times X_{h^\prime,0}^1(\Lambda)$. With little abuse of notation, we assume that the characteristic mesh size of the 3D mesh $\mathcal{T}^{\Omega}_h$ and the one of the 1D mesh $\mathcal{T}^{\Lambda}_{h'}$ are different. Concerning the multiplier space, let $\mathcal{G}_h = \{K \in \mathcal{T}^{\Omega}_h: \, K\cap \Lambda \neq \emptyset\}$, be the set of the 3D elements that intersect $\Lambda$. Then we define $Q_h=\{{\ld}_h : {\ld}_h \in P^0(K) \, \forall K \in \mathcal{G}_h\}$. We notice that the multiplier functions are defined on the 3D elements. With little abuse of notation, we denote with $Q_h$ also the restriction to $\Lambda$ of the space of piecewise constant functions defined in 3D. As a result, we have $Q_h \subset L^2(\Lambda) \subset H^{-\frac12}(\Lambda)$.
With this choice of multipliers the problem is not inf-sup stable, therefore the idea is to add a stabilization term $s({\ld}_h, {\md}_h): \ Q_h \times Q_h \rightarrow \mathbb{R}$ to \eqref{eq:prob2_discrete} following the approach introduced in \cite{burman2014}. 
The objective of this section is to analyze the stabilized version of the 3D-1D-1D problem: find $[u_h, {\ud}_h] \in X_h$ and ${\ld}_h \in Q_h$ such that
\begin{multline}\label{eq:prob2_stabilized}
a([u_h, {\ud}_h], [v_h, {\vd}_h]) +
b([v_h, {\vd}_h], {\ld}_h)
+ b([u_h, {\ud}_h], {\md}_h)
- s({\ld}_h, {\md}_h) 
= c(v_h) + d({\md}_h)\\
  \quad \forall [v_h, {\vd}_h] \in X_h, \,
  \forall {\md}_h \in Q_h\,.
\end{multline}
The idea of the stabilization strategy proposed in \cite{burman2014}
is to identify a new multiplier space $Q_H$, which is never implemented in practice, such that inf-sup stability with $X_h$ holds true. Then, the stabilization operator is designed to control the distance between $Q_h$ and $Q_H$ trough the following inequality
\begin{equation*}
    \|{\md}_h - \pi_H {\md}_h\|_{Q_H} \leq C s_h({\md}_h, {\md}_h)\,,
\end{equation*}
being $\pi_H$ a suitable projection operator $Q_h \rightarrow Q_H$. Applying the results obtained in \cite{burman2014} we will show that the well posedness of problem \eqref{eq:prob2_stabilized} is governed by the following lemma.
\begin{lemma}[Lemma 2.3 of \cite{burman2014}]\label{lemma23burman}
\begin{enumerate}
    \item If the $b(\cdot,\cdot),\, X_h,Q_H \rightarrow \R$ is inf-sup stable.
    \item If the stabilization operator $s_h(\cdot,\cdot),\, Q_H,Q_H \rightarrow \R$ is such that
    \begin{equation*}
        \beta \|{\md}_h\|_{H^{-\frac12}(\Lambda)} \leq \sup\limits_{v_h \in X_h} \frac{b(v_h,{\md}_h}{\vertiii{v_h})} + s_h({\md}_h, {\md}_h), \quad \forall {\md}_h \in Q_h\,.
    \end{equation*}
    \item If for any $[v_h, {\vd}_h]$ there exists a function $\xi_h\in Q_h$ depending on $[v_h, {\vd}_h]$, namely $\xi_h=\xi_h([v_h, {\vd}_h])$, s.t.
\begin{gather}
\label{stab_coercivity}
a([v_h, {\vd}_h],[v_h, {\vd}_h] ) + b([v_h, {\vd}_h], \xi_h) \geq \alpha_\xi \vertiii{[v_h, {\vd}_h]}_{X_h}\,,
\\
\label{stab_stability}
(s(\xi_{h}, \xi_{h}))^{\frac 12} \leq c_s \vertiii{[v_h, {\vd}_h]}_{X_h}\,,
\end{gather}
being $\vertiii{[\cdot \, ,\, \cdot] }_{X_h}$ a suitable discrete norm. 
\end{enumerate}
Then, problem \eqref{eq:prob2_stabilized} admits a unique solution.
\end{lemma}
For the proof of this result we remand the reader to Lemma 2.3 of \cite{burman2014}. In the reminder of this section we show how to find a multiplier space $Q_H$ and a stabilization operator $s_h$ such that all the assumptions of lemma \ref{lemma23burman} are satisfied.

>>>>

\def\patch{\omega _j}
\subsection{$\mathcal{T}^{\Omega}_h$ non conforming to $\Gamma$ and $\Lambda$} \label{sec:unfit2}
We analyze now the case in which the elements of the 3D mesh $\mathcal{T}^{\Omega}_h$ do not conform with the surface $\Gamma$ nor with $\Lambda$. In this case, the 3D-1D-1D
formulation, namely Problem 3D-1D-1D, is more suitable. 
%{\color{red} Add more details, we can refer also to cutFEM (Burman, Massing etc.) explaining the limitation of that approach (network case for example).}\\
Therefore we focus on the analysis of this discretization approach for Problem 2 solely.

\subsubsection{Problem 3D-1D-1D} We consider for the solutions $u_h$ and ${\ud}_h$ the spaces $X_h = X_{h,0}^1(\Omega) \times X_{h^\prime,0}^1(\Lambda)$. With little abuse of notation, we assume that the characteristic mesh size of the 3D mesh $\mathcal{T}^{\Omega}_h$ and the one of the 1D mesh $\mathcal{T}^{\Lambda}_{h'}$ are different. Concerning the multiplier space, let $\mathcal{G}_h = \{K \in \mathcal{T}^{\Omega}_h: \, K\cap \Lambda \neq \emptyset\}$, be the set of the 3D elements that intersect $\Lambda$. Then we define $Q_h=\{{\ld}_h : {\ld}_h \in P^0(K) \, \forall K \in \mathcal{G}_h\}$. We notice that the multiplier functions are defined on the 3D elements. With little abuse of notation, we denote with $Q_h$ also the restriction to $\Lambda$ of the space of piecewise constant functions defined in 3D. As a result, we have $Q_h \subset L^2(\Lambda) \subset H^{-\frac12}(\Lambda)$.
With this choice of multipliers the problem is not inf-sup stable, therefore the idea is to add a stabilization term $s({\ld}_h, {\md}_h): \ Q_h \times Q_h \rightarrow \mathbb{R}$ to \eqref{eq:prob2_discrete} following the approach introduced in \cite{burman2014}. 
The objective of this section is to analyze the stabilized version of the 3D-1D-1D problem: find $[u_h, {\ud}_h] \in X_h$ and ${\ld}_h \in Q_h$ such that
\begin{multline}\label{eq:prob2_stabilized}
a([u_h, {\ud}_h], [v_h, {\vd}_h]) +
b([v_h, {\vd}_h], {\ld}_h)
+ b([u_h, {\ud}_h], {\md}_h)
- s({\ld}_h, {\md}_h) 
= c(v_h) + d({\md}_h)\\
  \quad \forall [v_h, {\vd}_h] \in X_h, \,
  \forall {\md}_h \in Q_h\,.
\end{multline}
The idea of the stabilization strategy proposed in \cite{burman2014}
is to identify a new multiplier space $Q_H$, which is never implemented in practice, such that inf-sup stability with $X_h$ holds true. Then, the stabilization operator is designed to control the distance between $Q_h$ and $Q_H$ trough the following inequality
\begin{equation*}
    \|{\md}_h - \pi_H {\md}_h\|_{Q_H} \leq C s_h({\md}_h, {\md}_h)\,,
\end{equation*}
being $\pi_H$ a suitable projection operator $Q_h \rightarrow Q_H$. Applying the results obtained in \cite{burman2014} we will show that the well posedness of problem \eqref{eq:prob2_stabilized} is governed by the following lemma.
\begin{lemma}[Lemma 2.3 of \cite{burman2014}]\label{lemma23burman}
\begin{enumerate}
    \item If the $b(\cdot,\cdot),\, X_h,Q_H \rightarrow \R$ is inf-sup stable.
    \item If the stabilization operator $s_h(\cdot,\cdot),\, Q_H,Q_H \rightarrow \R$ is such that
    \begin{equation*}
        \beta \|{\md}_h\|_{H^{-\frac12}(\Lambda)} \leq \sup\limits_{v_h \in X_h} \frac{b(v_h,{\md}_h}{\vertiii{v_h})} + s_h({\md}_h, {\md}_h), \quad \forall {\md}_h \in Q_h\,.
    \end{equation*}
    \item If for any $[v_h, {\vd}_h]$ there exists a function $\xi_h\in Q_h$ depending on $[v_h, {\vd}_h]$, namely $\xi_h=\xi_h([v_h, {\vd}_h])$, s.t.
\begin{gather}
\label{stab_coercivity}
a([v_h, {\vd}_h],[v_h, {\vd}_h] ) + b([v_h, {\vd}_h], \xi_h) \geq \alpha_\xi \vertiii{[v_h, {\vd}_h]}_{X_h}\,,
\\
\label{stab_stability}
(s(\xi_{h}, \xi_{h}))^{\frac 12} \leq c_s \vertiii{[v_h, {\vd}_h]}_{X_h}\,,
\end{gather}
being $\vertiii{[\cdot \, ,\, \cdot] }_{X_h}$ a suitable discrete norm. 
\end{enumerate}
Then, problem \eqref{eq:prob2_stabilized} admits a unique solution.
\end{lemma}
For the proof of this result we remind the reader to Lemma 2.3 of \cite{burman2014}. In the remainder of this section we show how to find a multiplier space $Q_H$ and a stabilization operator $s_h$ such that all the assumptions of lemma \ref{lemma23burman} are satisfied.
